---
title: "An introduction to 'castarter' - content analysis starter toolkit for R"
author: "Giorgio Comai"
date: "`r Sys.Date()`"
output: github_document
bibliography: readme.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r knitr settings, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "data/readme/README-"
)
knitr::opts_knit$set(root.dir = normalizePath('./data'))
```

## Preliminary note (2018-01-14)

The new `castarter` aims to:

- make it possible to run (almost) all of its functions from a web interface (through a Shiny app), allowing to create textual datasets, export them, and conduct basic word frequency analysis virtually without knowing anything about R
- store parameters used for dataset creation in a format that is machine readbale, making it easy to share the procedure, and allow other users to replicate it
- automatically generate human-readable log files that detail the procedure
- have as few dependencies as possible on other packages
- reducing RAM requirements in order to allow for smooth analysis of relatively big datasets (e.g. >1 million items)
- facilitate integration with SQL databases

## Examples of output
Some examples of analysis of media contents conducted with `castarter` are available on the author's blog:

* [Word frequency of ‘Ukraine’, ‘Crimea’, and ‘Syria’ on Russia’s First Channel](http://www.giorgiocomai.eu/2015/11/03/word-frequency-of-ukraine-crimea-and-syria-on-russias-first-channel/)
* [Presidents and prime ministers in Romanian media](http://www.giorgiocomai.eu/2016/06/29/presidents-and-prime-ministers-in-romanian-media/)

Interactive examples:

* [Armenia's president website](https://giocomai.shinyapps.io/ArmeniaPresident/)
* [Official websites of South Caucasus presidents](https://giocomai.shinyapps.io/SouthCaucasusPresidents/)
* [Kremlin's official website - only time series](https://giocomai.shinyapps.io/kremlinregex/)

## Installation

N.B. Please consider that this document illustrates only the most basic parameters of `castarter` functions. More advanced options can be seen by running `?nameOfFunction`.

`castarter` is hosted on GitHub. The easiest way to install it on your system, is to run the following commands in R:

``` {r Install castarter, eval = FALSE}
install.packages("devtools")
devtools::install_github("giocomai/castarter")
```

For detailed install instructions, including references to packages that need to be preliminary installed on Linux system, as well as advanced setups (including Docker images), see [install instructions](./install.md).

Load `castarter` with:
```{r Load castarter, message=FALSE, warning=FALSE}
library("castarter")
```

## Using castarter
`castarter` facilitates downloading and extracting textual contents from a given section of a website. `castarter` can be used to analyse one single website, or more websites as part of a project, facilitating comparison among cases. In order to smoothen replication and prevent data losses, by default `castarter` stores a copy of downloaded web pages in html format. To do so, it includes a function that creates the relevant folder structure, and another function that facilitates archiving the html files in compressed format.

In `castarter` every dataset is expected to be extracted from a website, and each website is expected to be part of a project. This is useful to give a consistent way of storing the materials gathered and to facilitate comparison among cases, but it does not imply additional limitations: it is possible to have projects composed of a single website, and it is possible to make comparisons among websites that are part of different projects. 

[Readme to be completed]

## Disclaimer
`castarter` is under active development:

* some functions may not work as expected;
* documentation is incomplete (however, most paramaters should be self-descriptive);
* the code has not been optimised. 

## About the author
Giorgio Comai is a researcher at OBCT/CCI (https://www.balcanicaucaso.org/) and a data cruncher at EDJNET (https://www.europeandatajournalism.eu/). He has completed his PhD at the School of Law and Government, Dublin City University. You can find out more about his work on his [personal website](http://giorgiocomai.eu/). He is on Twitter: https://twitter.com/giocomai.
