---
title: "Install `castarter`"
author: "Giorgio Comai"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Install `castarter`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Preliminary steps
In order to use `castarter`, you need to install R and, if on Linux, a few dependencies. 

### On Fedora
```{install on Fedora}
sudo dnf install R R-RCurl curl-devel R-zoo R-XML openssl-devel libxml2-devel cairo-devel
```
### On Ubuntu
```{install on Ubuntu}
sudo apt-get install r-base r-cran-rcurl r-cran-xml libssl-dev libcairo-dev libcurl4-openssl-dev r-cran-slam libxml2-dev
```
### On Windows
Download the latest version of R from [R's official website]( https://www.r-project.org/{https://www.r-project.org/)

### Installing a user interface for R
No matter which platform you are using, it is suggested you install a dedicated user interface for R such as [RStudio](https://www.rstudio.com/products/rstudio/download/) - open source and available for Linux, Windows and Mac. 

## Installing 'castarter'
To actually install `castarter`, you can run these commands inside R: they will install the `devtoools` package, which allows to install `castarter` directly from Github.

```{r installing castarter, eval = FALSE}
install.packages("devtools")
devtools::install_github("giocomai/castarter")
```
